---
title: "JSON Vignette"
author: "Yinzhou Zhu"
date: "5/31/2020"
output: 
    rmarkdown::github_document:
        toc: true
        toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(httr)
library(jsonlite)
library(knitr)
```

## About JSON data
**JavaScript** is the most widely used language that powers interactions between client devices and server/web applications, where it allows programmers to implement complex features<sup>1</sup>. As a high-level language, it is also light-weight<sup>2</sup>.  
JavaScript Object Notation (**JSON**), as its name also implies, is a type of data that is easily converted from and to **JavaScript** objects<sup>3</sup>. **JSON** is stored as text string, which makes it possible to store **JavaScript** objects as text. It is now a very standard way for data communication over the network, which used to be primarily with **XML** format that is much heavier in weight.  
As described, **JSON** was built with purpose of being a standardized transfer language that is easy for human to read and at the same time easy for machine to parse and execute <sup>4</sup>.  
The flexibility of **JSON** means it's also suitable for general data transfer beyond **JavaScript** objects between devices across the web. This makes it a popular way for accessing data through cloud where **JSON API** is implemented<sup>5</sup>.  

## Packages for reading JSON  
### `rjson`, `jsonlite`, `RJSONIO` and `tidyjson`  
`rjson`, `jsonlite` and `RJSONIO` are 3 popular R packages used for working with **JSON** data. Functionality wise, they are all very similar, where they allow conversion of **R** objects from and to **JSON** data. The major differences being mainly: usage syntax, internal implementation methods (such as how data is read etc., thus affecting speed) and extended features. For example, `jsonlite` is able to stream to/from JSON file if the data that we are dealing with is large, which could be an advantage<sup>6</sup>. `tidyjson` on the other hand seems to be gaining traction. Under the same framework of `tidyverse`, this package could be particularly useful in streamlining our pipeline when appropriate<sup>7</sup>.  

### Why choosing `jsonlite`?  
I have chosen to work with `jsonlite` for the following reasons:  
* I have gotten most familiar with this package due to course usage.  
* It has advantage and flexibility for dealing with a large amount of data as mentioned above.  
* The overall consensus I got from reading online is that `jsonlite` strikes a very good balance between features and performance and is particulate in format conversions.  
  
## JSON example: NHL API
```{r NHL api}
api_url <- 'https://records.nhl.com/site/api'
nhl_franchise <- function(){
  franchises <- httr::GET(paste0(api_url, '/franchise'))
  franchises <- httr::content(franchises, 'text')
  franchises <- fromJSON(franchises, flatten = TRUE)
}
nhl_franchise_team_totals <- function(){
  franchises <- httr::GET(paste0(api_url, '/franchise-team-totals'))
  franchises <- httr::content(franchises, 'text')
  franchises <- fromJSON(franchises, flatten = TRUE)
}
nhl_franchise_records <- function(franchise_id, records='season'){
  if (!records%in%c('season', 'goalie', 'skater')) {
    stop("Records can only be 'season', 'goalie' or 'skater'.")
  }
  if (is.na(as.numeric(franchise_id))) {
    stop("Please enter a valid franchasie ID.")
  }
  if (records=='season') {  
    franchises <- httr::GET(paste0(api_url, '/franchise-season-records?cayenneExp=franchiseId=',franchise_id))
  } else if (records=='goalie') {
    franchises <- httr::GET(paste0(api_url, '/franchise-goalie-records?cayenneExp=franchiseId=',franchise_id))
  } else {
    franchises <- httr::GET(paste0(api_url, '/franchise-skater-records?cayenneExp=franchiseId=',franchise_id))
  }
  franchises <- httr::content(franchises, 'text')
  franchises <- fromJSON(franchises, flatten = TRUE)
}
```

## Formatted data
List of all franchises.
```{r list all franchises}
nhl_franchise()$data %>% kable()
```

Basic overview for New York Rangers and Boston Bruins.  
There seems to be more than one game type, but I summarized them all together here.  
```{r list NYR and BB}
totals <- nhl_franchise_team_totals()
nyr_bb_seasons <- totals$data %>% filter(franchiseId==6|franchiseId==10) %>% select(franchiseId, gamesPlayed, wins)
nyr_bb_games <- nyr_bb_seasons %>% group_by(franchiseId) %>% summarise(games_played=sum(gamesPlayed), wins=sum(wins))
nyr_bb_games <- nyr_bb_games %>% mutate(win_ratio=wins/games_played)
nyr_bb_games$wins<-NULL
rownames(nyr_bb_games)<-c('New York Rangers', 'Boston Ruins')
colnames(nyr_bb_games)<-c('Franchise ID', 'Games Played', 'Win Ratio')
kable(cbind(nyr_bb_games))
```

Whats the relationship between penalty minutes and wins ratio?  

```{r pentalty vs win_ratio}
selected_columns <- totals$data%>%group_by(gameTypeId)%>%select(triCode,firstSeasonId, gamesPlayed, gameTypeId, penaltyMinutes, wins) %>% mutate(win_ratio=wins/gamesPlayed, .keep = 'unused')
selected_columns <- selected_columns %>% mutate(game_type=paste0('Game Type ID: ', gameTypeId), .keep = 'unused') %>% mutate(era = if_else(firstSeasonId <=19911992,'Pre-Modern','Modern'), .keep = 'unused')
scatter_plot <-selected_columns %>% ggplot(aes(penaltyMinutes, win_ratio)) + geom_point(aes(color=penaltyMinutes)) + facet_wrap(vars(game_type), scales = "free") + xlab('Accumulated Penalty Time (Minutes)') + ylab('Accumulated Win Ratio') + geom_text(aes(label = triCode), color="red", data = subset(selected_columns, win_ratio>0.5), size = 2, nudge_y = 0.02)
scatter_plot + geom_smooth(method = lm) + theme(panel.background = element_rect(fill = 'transparent', color = 'black'), panel.grid = element_blank())
scatter_plot + geom_smooth() + theme(panel.background = element_rect(fill = 'transparent', color = 'black'), panel.grid = element_blank())
```

```{r pre-modern teams vs modern teams}
bar_plot <- selected_columns %>% ggplot(aes(era)) + geom_bar(aes(era)) + geom_bar(aes(fill = era), data = subset(selected_columns, win_ratio>0.4)) + facet_wrap(vars(game_type), scales = "free") + xlab("Era (Teams Joint Before/After 1991)") + ylab("Number of Teams")
bar_plot + theme(panel.background = element_rect(fill = 'transparent', color = 'black'), panel.grid = element_blank()) + scale_fill_discrete(name="Category", labels=c("Modern: Win Ratio > 0.3", "Pre-Modern: Win Ratio > 0.3"))
```

```{r road losses}
selected_columns <- totals$data%>%group_by(gameTypeId)%>%select(gamesPlayed, gameTypeId, triCode, roadLosses, roadWins, gameTypeId, wins) %>% mutate(win_ratio=wins/gamesPlayed, .keep = 'unused') %>% mutate(category_col= if_else(win_ratio<0.3,'T4', if_else(win_ratio>=0.3&win_ratio<0.4, 'T3', if_else(win_ratio>=0.4&win_ratio<0.5, 'T2', if_else(win_ratio>0.5, 'T1', 'NA')))))
selected_columns <- selected_columns %>% mutate(game_type=paste0('Game Type ID: ', gameTypeId), .keep = 'unused') %>% mutate(road_win_loss_ratio=roadWins/roadLosses, .keep = 'unused')
box_plot <- selected_columns %>% ggplot(aes(category_col, road_win_loss_ratio)) + geom_boxplot() + geom_jitter(aes(color=category_col)) + facet_wrap(vars(game_type), scales = 'free') + xlab("Team Tiers (Based on Total Win Ratio)") + ylab("Road Win-Loss Ratio")
box_plot + theme(panel.background = element_rect(fill = 'transparent', color = 'black'), panel.grid = element_blank(), legend.position = "none")
```


## References  
**1.** https://www.bigcommerce.com/ecommerce-answers/what-javascript-and-why-it-important/  
**2.** https://developer.mozilla.org/en-US/docs/Web/JavaScript  
**3.** https://www.w3schools.com/js/js_json_intro.asp  
**4.** https://www.json.org/json-en.html  
**5.** https://nordicapis.com/the-benefits-of-using-json-api/  
**6.** http://anotherpeak.org/blog/tech/2016/03/10/understand_json_3.html
**7.** https://cran.csiro.au/web/packages/tidyjson/vignettes/introduction-to-tidyjson.html
